EvoSwarm: An Event-based Architecture for Multi-population Optimization Algorithms

### 2020-05-12 draft.tex
Cloud experiment setup

\lstset{
basicstyle=\small\ttfamily,
}
\graphicspath{ {./images/} }

@@

Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.
%

@@

% Section headers should be headlines. Like "Measuring scalability in
% heterogenous systems", or something like that. Experiment 1 says
% nothing about what's inside (this also applies below) - JJ
\section{Measuring Scalability}
\label{sec:exp1}
% Experiment 1
As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
                                % reference - JJ
                                % done - Mario
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In a population-based algorithm like
EvoSwarm, several parameters can increase or decrease the workload of workers:

@@

Another additional issue is related to the algorithm itself. Many more
workers than populations will have a wider range of states of
evolution in the experiment; populations that will be mixed will have
very different fitness states, which means that a part of them will
probably have been not so useful, following the intermediate
disturbance hypothesis. When the number of workers is increased, the
difference between the state of evolution of populations will be
intermediate, leading to better results; eventually, when there are as
many workers as populations, the state will be quite similar, which is
shown with a smaller gap in the time won adding more workers.
%
% Boxplot of the total number of evaluations for 30 instances of function f4 , with
% different number of workers and populations sizes

%
%
% The effect of a
% different number of populations on execution time is reported in  Section
% \ref{sec:exp1results}.
% The results of the performance comparison between homogeneous and
% heterogeneous search strategies are presented in Section \ref{sec:exp2results}.     
%
%

%
%
%\begin{figure}[h!tbp]
%%    \centering
%    \includegraphics[width=\textwidth]{evalrate}
%%    \caption{Boxplot of the number of evaluations per second, for 30 instances of function $f_4$, with different number of workers and populations sizes }
%    \label{fig:spworker:ratio}
%\end{figure}

@@

% You should start writing this out soon... Don't leave it this way - JJ
Several options:
    A virtual machine with a container-optimized operating system, for example,
     Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
      virtual machine and run docker-compose commands as we would do in a local 
      workstation. % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ

    A Kubernetes cluster, this provides auto-scaling and management of containerized
    application, but it needs additional configuration for provisioning. All major
    cloud-providers have Kubernetes services.

    A container orchestration service, these services are compatible with docker-compose
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.

    % This is the second option. The first option was already tested, right? - JJ
As a proof-of-concept, we deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration (from deploying it locally using docker-compose)
and it is has a lower cost than a Kubernetes alternative.

@@

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. Multicontainer applications are
% Maybe an example - JJ
deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. When using the Fargate service as a backend, tasks are limited to a
maximum of 10 containers.

We deployed an AWS Fargate cluster and defined a Task configuration file with
8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
ten container limit, we only used seven workers and tested the same function as
in the previous experiment, but save resources only for dimensions 10 and 20.

The results ..

>>>>

\lstset{
basicstyle=\small\ttfamily,
}

\graphicspath{ {./images/} }

@@

Next, we specify a list indicating which of the 24 benchmark
functions will be tested. In the example, the experiment will use the first five
functions.  The {\em Instances} parameter indicates how many instances of each
function will be tested. {\em Instances}  has a default value of 15. We used this
value because it is the standard for the BBOB benchmark \cite{hansen2016coco}.
In the {\em Dimensions} parameter, we define a list of the dimensions that will
be tested, and we must select additional parameters for each dimension.
According to the maximum number of function evaluations (FEs) we mentioned
earlier, we define for each dimension: the number of populations that will be
generated in the setup, and for each population,  the number of generations, and
population size. Finally, the {\em Iterations} parameter indicates how many
complete loops will be performed. All these parameters give us the maximum
number of FEs that will be performed. For instance, for $D = 2$, the maximum
number of FEs is $200,000$, which is the same as $40*50*10*10$.

@@

\section{Measuring Scalability}
\label{sec:exp1}
% Experiment 1
As we mentioned before, reactive systems can scale by adding additional copies
of serverless functions. In our case, we can start additional worker containers
to have the same effect. Other authors, like Salza et al. \cite{salza2019speed},   % add
                                % reference - JJ
                                % done - Mario
have used a
similar architecture, which implies that worker nodes need to have at least a certain level
of complexity, in terms of execution time,  to effectively scale on multiple
nodes, tending to linear scalability. In a population-based algorithm like
EvoSwarm, several parameters can increase or decrease the workload of workers:

@@

Another additional issue is related to the algorithm itself. Many more
workers than populations will have a wider range of states of
evolution in the experiment; populations that will be mixed will have
very different fitness states, which means that a part of them will
probably have been not so useful, following the intermediate
disturbance hypothesis. When the number of workers is increased, the
difference between the state of evolution of populations will be
intermediate, leading to better results; eventually, when there are as
many workers as populations, the state will be quite similar, which is
shown with a smaller gap in the time won adding more workers.

@@

Container-based applications describe the deployment of isolated processes in a
consistent and portable way, going from development, testing, and production
independently on the kind of infrastructure. The term cloud-native describes
these container-based environments because they are deployed on an
infrastructure that abstracts the underlying compute, storage, and networking
primitives. This infrastructure does need to be a cloud provider; it could be a
developer's laptop, a single server, or a production environment. When we deploy
a cloud-native application to a cloud provider, we have more options to choose:

\begin{description}

  \item[A virtual machine] with a container-optimized operating system, for example,
  Google Compute Engine VMs, or Amazon EC2.  We still need to provision the
  virtual machine and run docker-compose commands as we would do in a local 
  workstation. Salza et al. \cite{salza2019speed} tested this option by using
  CoreOs cluster. 
  % Remember that Salza et al. used CoreOS, which is optimized for containers - JJ

  \item[A Kubernetes cluster] provides auto-scaling and management of containerized
  applications, but it needs additional configuration for provisioning. All major
  cloud-providers have Kubernetes services.

  \item[A container orchestration service] these services are compatible with docker-compose
    files for configuration and provide transparent provisioning. Examples of these
    services are Amazon Elastic Container Service,  Red Hat Openshift, and Marantis Docker Enterprise.

\end{description}

% This is the second option. The first option was already tested, right? - JJ
% This is the third option.  - M

As a proof-of-concept, we chose the third option and deployed the EvoSwarm application in Amazon Elastic
Container Service (ECS),  since it needs fewer changes in configuration 
(from deploying it locally using docker-compose) and it is has a lower cost than a Kubernetes alternative.

@@

Multicontainer applications are 
% Maybe an example - JJ
% This is shown next - Mario
deployed to ECS as tasks, and a single cluster can run multiple tasks in
parallel. When using the Fargate service as a backend, tasks are limited to a
maximum of 10 containers.

Using the ECS CLI, we can provision a cluster with a configuration file
specifying the computing resources needed. We can specify the configuration a Task 
by using a task definition file (see Listing~\ref{code:ecs}) in YAML format.
The main attributes are the {\texttt task\_size} indicating the memory and CPU units
assigned, and network configuration. The complete configuration of the task is generated
by the {\texttt compose command} that uses a {\texttt docker-compose.yml} to generate 
the configuration on-the-fly. For example to deployed the controller, setup and redis 
containers we used the file in Listing ~\ref{code:compose}.

We deployed two configurations, one that included all containers in a single Task,
and another one, where we assigned each worker to a single Task and the controller, 
redis, and setup in another independent Task.

For the first option, we deployed an AWS Fargate cluster and defined a Task configuration file with
8GBs of memory and 4096 computing units, equivalent to 4 CPUs. Because of the
ten container limit, we only used seven workers and tested function $f_4$ as
in the previous experiment, and to save resources we only tested for dimensions 10 and 20.

For the second option we assigned each worker to a single task with 4GBs of memory and 2048 computing units,
and scale the number of worker Tasks to eight.

\begin{lstlisting}[caption=ecs-params.yml , label=code:ecs]
version: 1
task_definition:
  task_execution_role: ecsTaskExecutionRole
  ecs_network_mode: awsvpc
  task_size:
    mem_limit: 8GB
    cpu_limit: 4096
run_params:
  network_configuration:
    awsvpc_configuration:
      subnets:
        - "subnet-0ca2a542bbd74239e"
        - "subnet-0dfd367b5ca531b07"
      security_groups:
        - "sg-0681c18469874b391"
      assign_public_ip: ENABLED
\end{lstlisting}

\begin{lstlisting}[caption={docker-compose.yml for setup, controller and redis containers}, label=code:compose]
  version: '3'
  services:
    redis:
      image: "redis:alpine"
      ports:
        - "6379:6379"
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: redis

    setup:
      image: mariosky/setup
      command: python experiment.py
      volumes:
        - /code
      environment:
        PYTHONUNBUFFERED: 1
        REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: setup

    controller:
      image: mariosky/controller
      volumes:
          - /code
      command: python controller.py
      environment:
          PYTHONUNBUFFERED: 1
          REDIS_HOST: localhost
      logging:
        driver: awslogs
        options:
          awslogs-region: us-west-2
          awslogs-group: docker-logs
          awslogs-stream-prefix: controller
  \end{lstlisting}

The results are shown in Table X.
